---
title: "blackjack-intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blackjackR-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The `blackjack` package provides a flexible and extensible framework for simulating the card game of Blackjack in R. Unlike many other implementations, this package supports **both single-player and multiplayer modes**, with integrated dealer logic and **automated strategy simulation**. It also leverages the {vctrs} package to enhance game by allowing users to **inspect and manipulate card properties** — such as rank, suit, and value—during play. 

Whether you're building bots, testing strategies, or just playing for fun, `blackjack` package offers a fully featured and customised experience! 

## Basic Blackjack Rules 

**General Rules** 

-   The goal is to get as close to 21 as possible without going over.
-   All number cards count as their face value; face cards (J, Q, K) are worth 10. Aces can count as 1 or 11, whichever benefits the hand more.
-   A Blackjack is a hand with an Ace + a 10-point card on the first deal.
-   If both player and dealer have Blackjack, the game is a push (tie); If a player busts (goes over 21), they automatically lose.

**Player Rules** 

-   Players are dealt two cards face up and can choose to: Hit (draw another card); Stand (keep their current hand); Double down (double their bet and receive only one more card)
-   Players can continue to hit until they stand or bust. 
-   In multiplayer mode, each player acts in turn.

**Dealer Rules** 

-   The dealer reveals one face-up card and keeps one face-down.
-   Dealer must hit until their hand totals 17 or more. Dealer must stand on all 17s, including soft 17 (Ace + 6), unless using smart logic.
-   If the dealer busts, all remaining players win; If the dealer does not bust, the hand is compared to each player’s hand to determine the outcome (win/lose/push).


## Function Examples 

Initialise the game by loading the library: 

```{r}
library(blackjack)
```

The package includes the functions as following to facilitate the game:

### 1. `create_shuffled_deck()`

Creates a shuffled Blackjack shoe (default = 208 cards). Checks the number of cards in the shuffled deck and inspects first `n` number of card: 

```{r}
# Create a shuffled deck of cards
deck <- create_shuffled_deck() 

# Check how many cards are in the deck
length(deck)

# Show the first 5 cards in the deck
head(deck, 5) 
```


### 2. `deal_hand()`

Deals cards from the deck, shows the cards dealt, and updates it:  

```{r}
# Deal 2 cards from the deck
deal <- deal_hand(deck, 2)  

# Show the cards dealt
deal$hand 

# Update the deck to remove dealt cards
deck <- deal$deck 

# Check how many cards are left in the deck
length(deck) 
```


### 3. `card_value()`

Calculates the total value of a hand, adjusting the Ace from 11 to 1 if needed to avoid busting: 

```{r}
# Create a hand (example: with an Ace and an 8)
hand <- c("A♣", "8♦")

# Get the total value of the hand (example: the result should be 19 from the example)
card_value(hand) 

# Get the value of a hand (example: with Ace downgraded to 1, 8, and 5, the result should be 14)
card_value(c("A♣", "8♦", "5♠")) 
```


### 4. `player_turn()`

Simulates a player's turn, allowing them to choose whether to hit or stand. In this example, we simulate the turn that always chooses to stand: 

```{r}
# Deal 2 cards to create a player hand
hand <- deal_hand(deck, 2)$hand
```

```{r}
# The game usually pauses for user input to decide each move.
# To automate the process, we provide a fake input function that always chooses "stand".
player_turn(hand, deck, input_fn = function(prompt) "s")
```


### 5. `auto_player_turn()`

Automatically draws cards until a threshold:

```{r}
# Deal 2 cards to create a player hand
hand <- deal_hand(deck, 2)$hand

# Simulate an automatic player turn with a threshold of 17
result <- auto_player_turn(hand, deck, threshold = 17)

# Show the final hand after the turn
result$hand

# Show the total value of the final hand
result$total
```


### 6. `dealer_turn()`

Simulates the dealer's turn, following standard rules:

```{r}
# Deal 2 cards to create the dealer's starting hand
hand <- deal_hand(deck, 2)$hand

# Simulate the dealer's turn
result <- dealer_turn(hand, deck)

# Show the dealer's final hand
result$hand

# Show the total value of the dealer's hand
result$total
```


### 7. `dealer_turn_smart()`

Simulates the dealer's turn with a smart strategy, considering the player's hand:

```{r}
# Deal 2 cards for the dealer's starting hand
dealer_hand <- deal_hand(deck, 2)$hand

# Set a fixed player hand for comparison
player_hand <- c("10♦", "8♣")

# Simulate the dealer's turn using a strategy that reacts to the player's hand
dealer_turn_smart(dealer_hand, player_hand, deck)
```


### 8. `play_blackjack()`

Simulates a full game of Blackjack involving both player and dealer turns. The game supports single or multiple players. In this example, we demonstrate a game with 1 player versus the dealer.

```{r}
# The game usually pauses for user input to decide each move.
# To automate the process, we provide a fake input function that always chooses "stand".
play_blackjack(input_fn = function(prompt) "s")
```


### 9. `play_blackjack_multi()`

Simulates a full round of Blackjack with multiple players competing against the dealer. The game supports an unlimited number of players. In this example, we demonstrate game with 2 player: 

```{r}
# The game usually pauses for user input to decide each move.
# To automate the process, we provide a fake input function that always chooses "stand".
play_blackjack(input_fn = function(prompt) "s")
```


### 10. `simulation_blackjack()`

Simulates multiple automated rounds of Blackjack: 

```{r}
# Simulate 100 rounds of Blackjack using a fixed player strategy (hit until reaching 16)
simulation_blackjack(threshold = 16, num_rounds = 100)
```

### 11. `create_suffle_deck()`

Create a full shuffled Blackjack shoe and inspect the first card’s properties:

```{r}
# Create a shuffled deck 
deck <- create_shuffled_deck()

# View the first card
deck[1]

# Check card properties
card_name(deck[1])     
card_rank(deck[1])    
card_symbol(deck[1])  
card_suit(deck[1])    
card_is_face(deck[1])  
card_value(deck[1])  
```


## Package Design Decisions 

For a detailed discussion of the internal design choices, integration with C++ and the {vctrs} package, as well as future development plans, please refer to the [Design](https://monasharp.github.io/game-package-bilbies/articles/blackjack-design.html) section.


